The \project{unRAVE} catalog

### 2016-09-05 article/unrave.tex
Minor edits before attacking discussion

We chose to adopt a data-driven model --- rather than a physics-based model ---
for this analysis.  Specifically, we will use an implementation of \thecannon\
\citep{Ness_2015,Ness_2016}.  Although this choice complicated the construction 
of our model (e.g., see Section \ref{sec:the-training-set}), a data-driven approach 
makes use of all available information in the spectrum and lowers the S/N ratio 
at which systematic effects begin to dominate.  In other words, in the low S/N
regime, a well-constructed data-driven model will yield more precise labels 
(e.g., stellar parameters and chemical abundances) than most physics-driven 
models\footnote{However, see \citet{Casey_2016a}}.  This is particularly relevant 
for the low-resolution \rave\ data analysed here, because about half of the 
spectra have S/N $\lesssim 50$.

@@

\begin{itemize}
\item We assume that any fibre- and time-dependent variations in spectral
resolution in the \rave\ spectra are negligible.
\item The \rave\ noise variances are approximately correct, independent between
pixels, and normally distributed.
\item We assume that the normalization procedure employed by the \rave\ pipeline
is invariant with respect to the labels we seek to measure (e.g., $\teff$, $\logg$,
or [Fe/H]), and invariant with respect to the S/N ratio.
\item We assume that stars with similar labels ($\teff$, $\logg$, and abundances)
have similar spectra.
\item A stellar spectrum is a smooth function of the label values for that star,
and we assume that the function is smooth enough within a sub-space of the labels
(e.g., the giant branch or the main-sequence) that it can be reasonably approximated 
with a low-order polynomial in label space.
\item The training set (Section~\ref{sec:the-training-set}) has mean accurate labels.
Note that we do not assume that \emph{every} label in the training is accurate -- we
can afford to have a small fraction of inaccurate labels (e.g., a few obvious 
misclassifications in the training set are affordable).
\item We assume the training set is representative of the test set (the \rave\ survey).
\end{itemize}

@@

\noindent{}where $y_{jn}$ is the pseudo-continuum-normalized flux for star $n$ at wavelength pixel
$j$, $\vecv(\ell_n)$ is the vectorizing function that takes as input the $K$ labels
$\ell_n$ for star $n$ and outputs functions of those labels as a vector of length
$D>K$, $\vectheta_j$ is a vector of length $D$ of parameters influencing the model at
wavelength pixel $j$, and $e_{jn}$ is the residual (noise).  Here we will only consider
vectorizing functions with polynomial expansions (e.g., $\teff^2$, see Sections 
\ref{sec:a-simple-model}-\ref{sec:evolved-star-model}).  The noise values $e_{jn}$ can 
be considered to be drawn from a Gaussian distribution with zero mean and variance 
$\sigma_{jn}^2 + s_j^2$, where $\sigma_{jn}^2$ is the variance in flux $y_{jn}$ and 
$s_j^2$ describes the excess variance at the $j$-th wavelength pixel.

At the \emph{training} step we fix the $K$-lists of labels for the $n$ training set.
At each wavelength pixel $j$, we then find the parameters $\vectheta_j$ and $s_j^2$
by optimizing the penalized likelihood function
\begin{eqnarray}\label{eq:train}
\vectheta_j,s^2_j &\leftarrow& \argmin{\vectheta,s}\left[
    \sum_{n=0}^{N-1} \frac{[y_{jn}-\vecv(\ell_n)\cdot\vectheta]^2}{\sigma^2_{jn}+s^2}
    + \sum_{n=0}^{N-1} \ln(\sigma^2_{jn}+s^2) + \Lambda{}Q(\vectheta)
    \right]
  \quad ,
\end{eqnarray}

\noindent{}where $\Lambda$ is a regularization parameter which we will heuristically set
in later sections, and $Q(\vectheta)$ is a regularizing function that encourages 
$\vectheta$ values to take on zero values without breaking convexity:
\begin{eqnarray}\label{eq:regularization-function}
	Q(\vectheta) = \sum_{d=1}^{D-1} |{\theta_d}| \quad .
\end{eqnarray}

@@

The \emph{test step} is where we fix the parameters $\vectheta_j,s_j^2$ at all wavelength
pixels $j$, and optimize the $K$-list of labels $\ell_m$ for each test set star $m$.  Here
the objective function is:
\begin{eqnarray}\label{eq:test}
  \ell_m &\leftarrow& \argmin{\ell}\left[
    \sum_{j=0}^{J-1} \frac{[y_{jm}-\vecv(\ell)\cdot\vectheta_j]^2}{\sigma_{jm}^2 + s_j^2}
    \right]
  \quad .
\end{eqnarray}

@@

We sought to construct a training set of stars across the main-sequence, the
sub-giant branch, and the red giant branch.  We demanded stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of O, Mg, Si, Ca, Al, Fe, and Ni.  This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth internal data release of the \ges\ survey includes 
giant and main-sequence stars, but only 142 overlap with \rave, which is too small to
be a useful training set for our purposes.  The thirteenth data release from the 
\project{Sloan Digital Sky Survey} \citep{sloan_dr13} includes labels for \apogee\ stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the \apogee\ main-sequence labels suffered from significant systematic effects.  
A flat, then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars incorrectly sit above an isochrone in a classical Hertzsprung-Russell diagram).
If we consider lower-resolution studies as potential training sets, there are 2,369
stars that overlap with \lamost\ --- of which 2,213 have positive S/N ratios in the 
$g$-band (\texttt{snrg}).  However, the labels are expectedly less precise given the
lower resolution, there are no elemental abundances available for the main-sequence 
stars\footnote{Abundance information is available for \lamost\ stars from \citet{Ho_2016},
but that sample contains only giant stars}, and the \lamost\ lower main-sequence suffers
from the same systematic effects seen in \apogee\ data.  The \galah\ survey has a good 
overlap with \rave, but results from that sample are not yet precise enough for our 
purposes.

These constraints forced us to construct a heterogeneous training set.  Given previous
successes in transferring high S/N ratio labels from \apogee\ \citep{Ness_2015,
Ness_2016,Ho_2016,Casey_2016b}, we chose to use the 1,355 stars in the \apogee---\rave\ 
overlap sample for giant star labels in the training set.  Of these, about 900 are 
giants according to \apogee.  From this sample we selected stars to have: metallicity
determinations ([Fe/H] $> -5$); S/N ratios of $>$200 in \apogee\ and $>$25 in \rave; and 
we further required that \aspcap\ did not report any peculiar flags 
(\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 536 stars along the giant 
branch, with metallicities ranging from $[{\rm Fe/H}] = -1.79$ to 0.26.  Intermediate 
tests with globular cluster members showed that the metallicity range of the training 
set needed to extend below $[{\rm Fe/H}] \lesssim -2$ in order for our catalog to be 
practically useful.  \stub{Supplemented this sample with known metal-poor stars in \rave}.

Assembling a suitable training set for the main-sequence and sub-giant branch was less
trivial.  There are no spectroscopic studies that extend the range of stellar types we 
are interested in (e.g., FGKM-type stars), and also have a large enough sample size that
overlaps with \rave.  Moreover, most of the spectroscopic studies we considered also 
showed a flat lower main-sequence, a systematic consequence of the analysis method adopted 
\citep[see][for discussion on this issue]{Bensby_2014}.  For these reasons we chose to make 
use of the \epic\ catalog \citep{Huber_2016} for the training set labels on the 
main-sequence and sub-giant branch.  The \epic\ catalog follows from the successful
\project{Kepler} input catalog \citep{Brown_2011}, and provides probabilistic stellar 
classifications for 138,600 stars in the \project{K2} fields based on the 
astrometric, asteroseismic, photometric, and spectroscopic information available for
every star.  There are 4,611 stars that overlap between \epic\ and \rave.

\epic\ differs from the \project{Kepler} input catalog because \epic\ does not 
benefit from having narrow-band $DDO_{51}$ photometry in order to aid dwarf/giant 
classification.  Despite this limitation the labels in the \epic\ catalog have 
already been shown to be accurate and trustworthy \citep{Huber_2016}.  However, 
when the posteriors are wide (i.e., the quoted confidence intervals are large) 
due to limited information available, it is possible that a star has been 
misclassified.  This is most prevalent for sub-giants, where \citet{Huber_2016} 
note that $\approx55-70$\% of sub-giants are misclassified as dwarfs.  The 
probability of misclassification is usually quantified in the uncertainties given
for each star; most dwarfs that have a higher likelihood of being sub-giants have
large confidence intervals.  Therefore, requiring low uncertainties will decrease 
the total sample size, but in practice it removes most misclassifications.  The 
situation is far more favourable for dwarfs and giants.  Only $1-4$\% of giant 
stars are misclassified as dwarfs, and about 7\% of dwarfs are misclassified as 
giants.  To summarise, the \epic\ labels with narrow confidence intervals are 
usually of high fidelity, and given that we have spectra, we can identify any
spurious misclassifications.

We sought to have a small overlap between our giant and main-sequence star training
sets.  Most of our giant training set is encapsulated within $0 < \logg < 3.5$, 
however there is a sparse sampling of stars reaching to $\logg \approx 4$.  We
required $\logg > 3.5$ for the \epic\ main-sequence/sub-giant star training set,
allowing for $\approx0.5$~dex of overlap between the two training sets.  We further
employed the following quality constraints: the upper and lower confidence intervals 
in $\teff$ must be below 150~K; the upper and lower confidence intervals in $\logg$ 
must be less than 0.15~dex; the $S/N$ of the \rave\ spectra must exceed 
30~pixel$^{-1}$; and $\teff \leqslant 6750$~K.  Unfortunately these strict constraints
removed most metal-poor stars, which we later found to cause the test labels to have
under-predicted abundances for dwarfs of low metallicity.  For this reason we relaxed
(ignored) those quality constraints for stars with $[{\rm Fe/H}] < -1$.  
\stub{Include known metal-poor stars in \rave?}. After training
a model based on main-sequence and giant stars (Section \ref{sec:the-model}), we found 
we could identify misclassifications by leave-one-out cross-validation.  However, we 
chose not to do this because the number of likely misclassifications in the training
set was negligible ($\approx1$\%), and the improvement in main-sequence test set labels
was minimal.  The distilled sample of the \rave--\epic\ overlap catalog contains 583 
stars (of 4,611).

@@

We used no regularization ($\Lambda = 0$) for this model.  After training the model we
treated all \Nspectra\ spectra as test set objects.  In the right-hand panel of Figure 
\ref{fig:tes-set-density} we show the effective temperature $\teff$ and surface gravity 
$\logg$ for all spectra.  The main-sequence and red giant branch are clearly visible.  
However, the details of stellar evolution are no longer present: the sub-giant branch is 
not discernible, and there are a number of systematic artefacts (over-densities) present
in label space.  These artefacts disappear when we require additional quality constraints 
(e.g., no peculiar morphological classifications), but the complexity of the 
Hertzsprung-Russell diagram is still not present.  Thus, we concluded that while this 
model could be useful for deriving stellar classifications (e.g., F2-type giant), the 
labels are too imprecise.

@@

We inferred labels for all \rave\ spectra using this main-sequence/sub-giant star model.  
The results for the survey sample are shown in the left panel of Figure \ref{fig:test-set-density}.  
The increased density of Solar-type stars is consistent with \rave\
observing stars in the local neighbourhood, and the high number of turn-off
and main-sequence stars relative to the sub-giant branch is expected from 
the relative lifetimes of these evolutionary phases.  An over-density
of stars near the giant branch clump is also present.  This artefact is due 
to having giant stars in the test set, but not in the training set, and the
model is (poorly) extrapolating outside the convex hull of the training set.

@@

Similar to Sections \ref{sec:a-simple-model} and \ref{sec:unevolved-star-model},
we used a quadratic vectorizer for the giant star model.  Here the terms are 
expanded in the same way as equation \ref{eq:vectorizer-three-label}, only
with nine labels instead of three.  We set the regularization hyperparameter $\Lambda$
in the same way described in Section \ref{sec:unevolved-star-model}, using the same 30 trials of $\Lambda$.
The results are shown in Figure \ref{fig:set-hyperparameters}, where
again the enveloped region represents the minimum and maximum change in RMS label
deviation with respect to the unregularized case.  At the point of maximum improvement 
near $\Lambda = X.XX$, the RMS in all nine labels has decreased by at least $X$\%, 
and up to $X$\%.  Here the regularization strength also produces a sparser matrix
of $\vectheta$, with $\approx20$\% more terms (mostly cross-terms) having zero-valued entries.
Based on the increased model sparsity and decreasing RMS deviation in the labels, 
we adopt $\Lambda = X.XX$ for the giant star model.

Results from cross-validation on our regularized model imply that the precision
in labels is XXX~K in $\teff$, X.XX~dex in $\logg$, and between X.XX~dex (EL) and 
X.XX~dex (EL) depending on the elemental abundance.  We inferred labels for all
\Nspectra\ \rave\ spectra using this model, and these results are summarized in
Figure \ref{fig:test-set-density} (middle panel).  The red clump is clearly visible
and in the expected location, without requiring any post-analysis calibration.
Artefacts due to dwarf stars in the test set are also present.

@@

We have derived labels for all \Nspectra\ \rave\ spectra using the three models
described in previous sections.  The results from the joint model --- that includes 
the main-sequence, sub-giant and red giant branch --- shows that a single quadratic
model is too simple for the \rave\ spectral range.  The other models have problems,
too: unrealistic over-densities in label space show that the main-sequence model and
the giant model make very poor extrapolations for stars outside their respective
training sets.  For these reasons we were forced to exclude or severely penalize 
incorrect results from both models.

@@

In Figure \ref{fig:joint-model-differences} we show the differences in effective temperature
$\teff$ and surface gravity $\logg$ between: the main-sequence model and the joint
model; and the differences between the red giant branch model and the joint model. We have scaled
the differences in $\teff$ and $\logg$ to make the central peak near $(0, 0)$ to 
% ARC to HOGG: This is a total hack. Should we fit for these and \rho_{\teff,\logg} or are you content with the hack?
be approximately isotropic by setting
	$\delta_{T_{\rm eff}} = 90$~K for the main-sequence model, 
	$\delta_{T_{\rm eff}} = 50$~K for the giant model, and 
	$\delta_{\log{g}} = 0.15$~dex for both models.
The stars within the peak at $(0, 0)$ represent objects where
the joint model and the comparison model both report similar labels.  The artefacts 
seen in the Hertszprung-Russell diagrams in Figure \ref{fig:test-set-density} are
also present in Figure \ref{fig:joint-model-differences} as over-densities far away
from the central peak.  Therefore, we can 
adopt the scaled distance in labels $\teff$ and $\logg$ from the joint model,
% ARC to HOGG: Please suggest any nomenclature or definition changes as you see fit!
\begin{eqnarray}
	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{joint}}{\delta_{\log{g},ms}}\right)^2   \nonumber \\
	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{joint}}{\delta_{\log{g},giant}}\right)^2   \quad ,
\end{eqnarray}

@@

We treated the stacked spectra as normal survey stars.  We inferred labels
using all models and joined the results as per Section \ref{sec:joining-the-models}.  The labels inferred from the stacked spectra served as the basis
of comparison for all individual visits to that object.  Figure 
\ref{fig:formal-errors-comparison} shows the difference in labels 
between the comparison spectrum and a repeat visit, normalized by their 
formal errors summed in quadrature (e.g., 
$\Delta\logg/\sqrt{\sigma_{\logg,stacked}^2 + \sigma_{\logg,visit}^2}$).
If our measurements were unbiased by S/N and the formal errors were 
representative, these values should be normally distributed with a zero 
mean and unit variance.
\stub{Are they? (Probably). How much should we inflate our formal errors, and should this be as a function of S/N? Note that in all Figures and tables,
we show these inflated errors only, as we gauge them to be more representative.}

@@

As an initial point of comparison, we cross-matched our results against the 
official fourth \rave\ data release (Figure \ref{fig:rave-dr4-comparison}).
In order to provide a fair comparison, we only show stars that meet a number
of quality flags in \emph{both} samples. Specifically we require that the 
$S/N$ (as measured by our analysis) exceeds 10~pixel$^{-1}$, the reduced
$\chi_r^2$ in our analysis $< 3$, and that the \texttt{QK} flag reported by
\citet{Kordopatis_2013} is zero, indicating no problems with quality control.
There is very good agreement in $\teff$, with a bias and RMS of just XX~K 
and XXX~K, respectively.  The offset in $\logg$ on the giant branch between this study and 
\citet{Kordopatis_2013} has been noted in other studies, and this has issue
has been resolved in the fifth \rave\ data release by correcting $\logg$
values with a small sample of asteroseismic targets.  There is also a
slight discrepancy in the $\log$ values along the main-sequence, where our
work tends to taper down towards higher $\logg$ values at cooler temperatures,
and the \rave\ DR4 sample tends to have a slightly flatter lower main-sequence.
This difference is not likely to have a very significant effect on the detailed
abundance or spectrophotometric distance determinations between these studies 
\citep{Binney_2014}.

@@

% What information have we extracted: almost maximal use of giant branch,
% lesss so for dwarfs.

>>>>

We chose to adopt a data-driven model --- rather than a physics-based model ---
for this analysis.  Specifically, we will use an implementation of \thecannon\
\citep{Ness_2015,Ness_2016}.  Although this choice complicated the construction 
of our model (e.g., see Section \ref{sec:the-training-set}), a data-driven approach 
makes use of all available information in the spectrum and lowers the S/N ratio 
at which systematic effects begin to dominate.  In other words, in the low S/N
regime, a well-constructed data-driven model will yield more precise \emph{labels}
(e.g., stellar parameters and chemical abundances) than most physics-driven 
models\footnote{However, see \citet{Casey_2016a}}.  This is particularly relevant 
for the low-resolution \rave\ data analysed here, because about half of the 
spectra have S/N $\lesssim 50$~pixel$^{-1}$.

@@

\begin{itemize}
\item We assume that any fibre- and time-dependent variations in spectral
resolution in the \rave\ spectra are negligible.
\item The \rave\ noise variances are approximately correct, independent between
pixels, and normally distributed.
\item We assume that the normalization procedure employed by the \rave\ pipeline
is invariant with respect to the labels we seek to measure (e.g., $\teff$, $\logg$,
or [Fe/H]), and invariant with respect to the S/N ratio.  In other words, we assume
that the normalization procedure does not produce different results for high S/N
spectra compared to low S/N spectra, nor does the normalization procedure vary 
non-linearly with respect to stellar parameters (e.g., [Fe/H]).
\item We assume that stars with similar labels ($\teff$, $\logg$, and abundances)
have similar spectra.
\item A stellar spectrum is a smooth function of the label values for that star,
and we assume that the function is smooth enough within a sub-space of the labels
(e.g., the giant branch or the main-sequence) that it can be reasonably approximated 
with a low-order polynomial in label space.
\item The training set (Section~\ref{sec:the-training-set}) has mean accurate labels.
Note that we do not assume that \emph{every} label in the training is accurate -- we
can afford to have a small fraction of inaccurate labels (e.g., a few obvious 
misclassifications in the training set are affordable).
\item We assume the training set is representative of the test set (the \rave\ survey).
\end{itemize}

@@

\noindent{}where $y_{jn}$ is the pseudo-continuum-normalized flux for star $n$ at wavelength pixel
$j$, $\vecv(\ell_n)$ is the vectorizing function that takes as input the $K$ labels
$\ell_n$ for star $n$ and outputs functions of those labels as a vector of length
$D>K$, $\vectheta_j$ is a vector of length $D$ of parameters influencing the model at
wavelength pixel $j$, and $e_{jn}$ is the residual (noise).  Here we will only consider
vectorizing functions with second-order polynomial expansions (e.g., $\teff^2$, see Sections 
\ref{sec:a-simple-model}-\ref{sec:evolved-star-model}).  The noise values $e_{jn}$ can 
be considered to be drawn from a Gaussian distribution with zero mean and variance 
$\sigma_{jn}^2 + s_j^2$, where $\sigma_{jn}^2$ is the variance in flux $y_{jn}$ and 
$s_j^2$ describes the excess variance at the $j$-th wavelength pixel.

At the \emph{training} step we fix the $K$-lists of labels for the $n$ training set stars.
At each wavelength pixel $j$, we then find the parameters $\vectheta_j$ and $s_j^2$
by optimizing the penalized likelihood function
\begin{eqnarray}\label{eq:train}
\vectheta_j,s^2_j &\leftarrow& \argmin{\vectheta,s}\left[
    \sum_{n=0}^{N-1} \frac{[y_{jn}-\vecv(\ell_n)\cdot\vectheta]^2}{\sigma^2_{jn}+s^2}
    + \sum_{n=0}^{N-1} \ln(\sigma^2_{jn}+s^2) + \Lambda{}Q(\vectheta)
    \right]
  \quad ,
\end{eqnarray}

\noindent{}where $\Lambda$ is a regularization parameter which we will heuristically set
in later sections, and $Q(\vectheta)$ is a L1 regularizing function that encourages 
$\vectheta$ values to take on zero values without breaking convexity:
\begin{eqnarray}\label{eq:regularization-function}
	Q(\vectheta) = \sum_{d=1}^{D-1} |{\theta_d}| \quad .
\end{eqnarray}

@@

The \emph{test step} is where we fix the parameters $\vectheta_j,s_j^2$ at all wavelength
pixels $j$, and optimize the $K$-list of labels $\ell_m$ for the $m$-th test set star.  Here
the objective function is:
\begin{eqnarray}\label{eq:test}
  \ell_m &\leftarrow& \argmin{\ell}\left[
    \sum_{j=0}^{J-1} \frac{[y_{jm}-\vecv(\ell)\cdot\vectheta_j]^2}{\sigma_{jm}^2 + s_j^2}
    \right]
  \quad .
\end{eqnarray}

@@

We sought to construct a training set of stars across the main-sequence, the
sub-giant branch, and the red giant branch.  We required stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of O, Mg, Si, Ca, Al, Fe, and Ni.  This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth internal data release of the \ges\ survey includes 
giant and main-sequence stars, but only 142 overlap with \rave, which is too small to
be a useful training set for our purposes.  The thirteenth data release from the 
\project{Sloan Digital Sky Survey} \citep{sloan_dr13} includes labels for \apogee\ stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the \apogee\ main-sequence labels suffered from significant systematic effects.  
A flat, then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars incorrectly sit above an isochrone in a classical Hertzsprung-Russell diagram).
If we consider lower-resolution studies as potential training sets, there are 2,369
stars that overlap with \lamost\ --- of which 2,213 have positive S/N ratios in the 
$g$-band (\texttt{snrg}).  However, the labels are expectedly less precise given the
lower resolution, there are no elemental abundances available for the main-sequence 
stars\footnote{Abundance information is available for \lamost\ stars from \citet{Ho_2016},
but that sample contains only giant stars}, and the \lamost\ lower main-sequence suffers
from the same systematic effects seen in \apogee\ data.  The \galah\ survey has a good 
overlap with \rave, but results from that sample are not yet precise enough for our 
purposes.

These constraints forced us to construct a heterogeneous training set.  Given previous
successes in transferring high S/N ratio labels from \apogee\ \citep{Ness_2015,
Ness_2016,Ho_2016,Casey_2016b}, we chose to use the 1,355 stars in the \apogee---\rave\ 
overlap sample for giant star labels in the training set.  Of these, about 900 are 
giants according to \apogee.  From this sample we selected stars to have: 
determinations in all abundances of interest ([X/H] $> -5$ for O, Mg, Al, Si, Ca, Fe, and Ni); S/N ratios of $>$200 in \apogee\ and $>$25 in \rave; and 
we further required that \aspcap\ did not report any peculiar flags 
(\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 536 stars along the giant 
branch, with metallicities ranging from $[{\rm Fe/H}] = -1.79$ to 0.26.  Intermediate 
tests with globular cluster members showed that the metallicity range of the training 
set needed to extend below $[{\rm Fe/H}] \lesssim -2$ in order for our catalog to be 
practically useful.  \stub{Supplemented this sample with known metal-poor stars in \rave}.

Assembling a suitable training set for the main-sequence and sub-giant branch was less
trivial.  There are no spectroscopic studies that extend the range of stellar types we 
are interested in (e.g., FGKM-type stars), and which have a large enough sample size 
overlapping with \rave.  Moreover, most of the spectroscopic studies we considered also 
showed a flat lower main-sequence, a systematic consequence of the analysis method adopted 
\citep[see][for discussion on this issue]{Bensby_2014}.  For these reasons we chose to make 
use of the \epic\ catalog \citep{Huber_2016} for the training set labels on the 
main-sequence and sub-giant branch.  The \epic\ catalog follows from the successful
\project{Kepler} input catalog \citep{Brown_2011}, and provides probabilistic stellar 
classifications for 138,600 stars in the \project{K2} fields based on the 
astrometric, asteroseismic, photometric, and spectroscopic information available for
every star.  There are 4,611 stars that overlap between \epic\ and \rave.

\epic\ differs from the \project{Kepler} input catalog because \epic\ does not 
benefit from having narrow-band $DDO_{51}$ photometry in order to aid dwarf/giant 
classification.  Despite this limitation the labels in the \epic\ catalog have 
already been shown to be accurate and trustworthy \citep{Huber_2016}.  However, 
when the posteriors are wide (i.e., the quoted confidence intervals are large) 
due to limited information available, it is possible that a star has been 
misclassified.  This is most prevalent for sub-giants, where \citet{Huber_2016} 
note that $\approx55-70$\% of sub-giants are misclassified as dwarfs.  The 
probability of misclassification is usually quantified in the uncertainties given
for each star; most dwarfs that have a higher possibility of being sub-giants have
large confidence intervals.  Therefore, requiring low uncertainties will decrease 
the total sample size, but in practice it removes most misclassifications.  The 
situation is far more favourable for dwarfs and giants.  Only $1-4$\% of giant 
stars are misclassified as dwarfs, and about 7\% of dwarfs are misclassified as 
giants.  To summarise, the \epic\ labels with narrow confidence intervals are 
usually of high fidelity, and given that we have spectra, we can identify any
spurious misclassifications.

We sought to have a small overlap between our giant and main-sequence star training
sets.  Most of our giant training set is encapsulated within $0 < \logg < 3.5$, 
however there is a sparse sampling of stars reaching to $\logg \approx 4$.  We
required $\logg > 3.5$ for the \epic\ main-sequence/sub-giant star training set,
allowing for $\approx0.5$~dex of overlap between the two training sets.  We further
employed the following quality constraints: the upper and lower confidence intervals 
in $\teff$ must be below 150~K; the upper and lower confidence intervals in $\logg$ 
must be less than 0.15~dex; the $S/N$ of the \rave\ spectra must exceed 
30~pixel$^{-1}$; and $\teff \leqslant 6750$~K.  Unfortunately these strict constraints
removed most metal-poor stars, which we later found to cause the test labels to have
under-predicted abundances for dwarfs of low metallicity.  For this reason we relaxed
(ignored) those quality constraints for stars with $[{\rm Fe/H}] < -1$.  
\stub{Include known metal-poor stars in \rave?}. After training
a model based on main-sequence and giant stars (Section \ref{sec:the-model}), we found 
we could identify misclassifications by leave-one-out cross-validation.  However, we 
chose not to do this because the number of likely misclassifications in the training
set was negligible ($\approx1$\%), and the improvement in main-sequence test set labels
was minimal.  The distilled sample of the \rave--\epic\ overlap catalog contains 583 
stars (of 4,611).  The full training set for each model (see next sections) is shown
in Figure~\ref{fig:training-set-hrd}.

@@

We used no regularization ($\Lambda = 0$) for this model.  After training the model we
treated all \Nspectra\ spectra as test set objects.  In the right-hand panel of Figure 
\ref{fig:test-set-density} we show the effective temperature $\teff$ and surface gravity 
$\logg$ for all spectra.  The main-sequence and red giant branch are clearly visible.  
However, the details of stellar evolution are no longer present: the sub-giant branch is 
not discernible, and there are a number of systematic artefacts (over-densities) present
in label space.  These artefacts disappear when we require additional quality constraints 
(e.g., no peculiar morphological classifications), but the complexity of the 
Hertzsprung-Russell diagram is still not present.  Thus, we concluded that while this 
model could be useful for deriving stellar classifications (e.g., F2-type giant), the 
labels are too imprecise.

@@

We inferred labels for all \rave\ spectra using this main-sequence/sub-giant star model.  
The results for the survey sample are shown in the centre panel of Figure \ref{fig:test-set-density}.  
The increased density of Solar-type stars is consistent with \rave\
observing stars in the local neighbourhood, and the high number of turn-off
and main-sequence stars relative to the sub-giant branch is expected from 
the relative lifetimes of these evolutionary phases.  An over-density
of stars near the base of the giant branch is also present.  This artefact is due 
to having giant stars in the test set, but not in the training set, and the
model is (poorly) extrapolating outside the convex hull of the training set.

@@

Similar to Sections \ref{sec:a-simple-model} and \ref{sec:unevolved-star-model},
we used a quadratic vectorizer for the giant star model.  Here the terms are 
expanded in the same way as equation \ref{eq:vectorizer-three-label}, only
with nine labels instead of three.  We set the regularization hyperparameter $\Lambda$
in the same way described in Section \ref{sec:unevolved-star-model}, using the same 30 trials of $\Lambda$.
The results are shown in Figure \ref{fig:set-hyperparameters}, where
again the enveloped region represents the minimum and maximum change in RMS label
deviation with respect to the unregularized case.  At the point of maximum improvement 
near $\Lambda = X.XX$, the RMS in all nine labels has decreased by up to $X$\%.  Here the regularization strength also produces a sparser matrix
of $\vectheta$, with $\approx20$\% more terms (mostly cross-terms) having zero-valued entries.
Based on the increased model sparsity and decreasing RMS deviation in the labels, 
we adopt $\Lambda = X.XX$ for the giant star model.

Results from cross-validation on our regularized model imply that the precision
in labels is XXX~K in $\teff$, X.XX~dex in $\logg$, and between X.XX~dex (EL) and 
X.XX~dex (EL) depending on the elemental abundance.  We inferred labels for all
\Nspectra\ \rave\ spectra using this model, and these results are summarized in
the right panel of Figure \ref{fig:test-set-density}.  The red clump is clearly visible
and in the expected location, without requiring any post-analysis calibration.
However, artefacts due to dwarf stars in the test set are also present.

@@

We have derived labels for all \Nspectra\ \rave\ spectra using the three models
described in previous sections.  The results from the joint model --- that includes 
the main-sequence, sub-giant and red giant branch --- shows that a single 3-label quadratic
model is too simple for the \rave\ spectral range.  The other models have problems,
too: unrealistic over-densities in label space show that the main-sequence model and
the giant model make very poor extrapolations for stars outside their respective
training sets.  For these reasons we were forced to exclude or severely penalize 
incorrect results from both models.

@@

In Figure \ref{fig:joint-model-differences} we show the differences in effective temperature
$\teff$ and surface gravity $\logg$ between: the main-sequence model and the joint
model; and the differences between the red giant branch model and the joint model. We have scaled
the differences in $\teff$ and $\logg$ to make the central peak near $(0, 0)$ to 
% ARC to HOGG: This is a total hack. Should we fit for these and \rho_{\teff,\logg} or are you content with the hack?
be approximately isotropic by setting
	$\delta_{T_{\rm eff}} = 90$~K for the main-sequence model, 
	$\delta_{T_{\rm eff}} = 50$~K for the giant model, and 
	$\delta_{\log{g}} = 0.15$~dex for both models.
The stars within the peak at $(0, 0)$ represent objects where
the joint model and the comparison model both report similar labels.  The artefacts 
seen in the Hertzsprung-Russell diagrams in Figure \ref{fig:test-set-density} are
also present in Figure \ref{fig:joint-model-differences} as over-densities far away
from the central peak.  Therefore, we can 
adopt the scaled distance in labels $\teff$ and $\logg$ from the joint model,
% ARC to HOGG: Please suggest any nomenclature or definition changes as you see fit!
\begin{eqnarray}
	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{joint}}{\delta_{\log{g},ms}}\right)^2   \nonumber \\
	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{joint}}{\delta_{\log{g},giant}}\right)^2   \quad ,
\end{eqnarray}

@@

We treated the stacked spectra as normal survey stars.  We inferred labels
using all models and joined the results as per Section \ref{sec:joining-the-models}.  The labels inferred from the stacked spectra served as the basis
of comparison for all individual visits to that object.  Figure 
\ref{fig:formal-errors-comparison} shows the difference in labels 
between the comparison spectrum and a repeat visit, normalized by their 
formal errors summed in quadrature (e.g., 
$\Delta\logg/\sqrt{\sigma_{\logg,stacked}^2 + \sigma_{\logg,visit}^2}$).
If our measurements were unbiased by S/N and the formal errors were 
representative, these values should be normally distributed with a zero 
mean and unit variance.
\stub{Are they? (Probably not!). How much should we inflate our formal errors, and should this be as a function of S/N? Note that in all Figures and tables,
we show these inflated errors only, as we gauge them to be more representative.}

@@

As an initial point of comparison, we cross-matched our results against the 
official fourth \rave\ data release (Figure \ref{fig:rave-dr4-comparison}).
In order to provide a fair comparison, we only show stars that meet a number
of quality flags in \emph{both} samples. Specifically we require that the 
$S/N$ (as measured by our analysis) exceeds 10~pixel$^{-1}$, the reduced
$\chi_r^2$ in our analysis $< 3$, and that the \texttt{QK} flag from
\citet{Kordopatis_2013} is zero, indicating that no problems were reported by the pipeline.
There is very good agreement in $\teff$, with a bias and RMS of just XX~K 
and XXX~K, respectively.  The offset in $\logg$ on the giant branch between this study and 
\citet{Kordopatis_2013} has been noted in other studies, and this has issue
has been resolved in the fifth \rave\ data release by correcting $\logg$
values with a small sample of asteroseismic targets.  There is also a
slight discrepancy in the $\log$ values along the main-sequence, where our
work tends to taper down towards higher $\logg$ values at cooler temperatures,
and the \rave\ DR4 sample tends to have a slightly flatter lower main-sequence.
This difference is not likely to have a very significant effect on the detailed
abundance or spectrophotometric distance determinations between these studies 
\citep{Binney_2014}.

@@

% What we have presented, an independent analysis that makes maximal use of giant branch
% less-so for dwarfs
% the intent is for usefulness with chemo-dynamic studies.

% Caveats in the results:
% MULTIPLE MODELS
% HETEROGENEOUS TRAINING SET
% METAL POOR ABUNDANCES

% UNSURE IF TWO ABUNDANCE SCALES ARE ON THE SAME SCALE --!
% even if they are on the same scale...
% metal-poor dwarfs/sub-giants
